DRAFT \\ From physical principles to classical and quantum \\ Hamiltonian and Lagrangian particle mechanics

### 2017-01-03 2015CompleteDerivation/completeDerivation.tex
Reviewing with Christine

For the tool set to be on par with classical mechanics we need to address three requirements. We need to be able to compare two states and ask what fraction of them share the same configuration: this will give us an inner product. We need to be able to express distributions and expectations for different quantities: this will give us self-adjoint linear operators. We then need to identify the self-adjoint linear operators for the quantities $k_i$ and $p_i$ and to be able express the distribution of material in terms of those. Let's start with the first.

The idea is that, given two states $\psi_1$ and $\psi_2$, we should be able to tell how much of $\psi_1$ is prepared in the same way as of $\psi_2$. That is $\psi_1 = \alpha \psi_2 + \psi_3$ is the combination of the component that is prepared like $\psi_2$ (e.g. same distributions in position and momentum) and a component that is prepared completely differently. Let $\mathsf{P}_{\psi_2} : \Psi \rightarrow \Psi$ be the operator that gives us the component $\mathsf{P}_{\psi_2} (\psi_1) \equiv \alpha \psi_2$ prepared like $\psi_2$. This operator is linear in $\psi_1$: the component of the combined system that is prepared like $\psi_2$ is the sum of the subcomponents that are prepared like $\psi_2$. It also gives us the same result if applied once or twice: once we found the component that is prepared like $\psi_2$, it remains the same. $\mathsf{P}_{\psi_2}$ is therefore a projection: a linear operator such that $\mathsf{P}_{\psi_2}(\mathsf{P}_{\psi_2} (\psi_1))= \mathsf{P}_{\psi_2} (\psi_1)$.

If $\psi_1$ is already prepared like $\psi_2$, then its projection on $\psi_2$ will be equal to $\psi_1$. Conversely, if the projection of $\psi_1$ is equal to itself, then it means that the state is already prepared like $\psi_2$. That is: $\mathsf{P}_{\psi_2} (\psi_1) = \psi_1 \Leftrightarrow \psi_1 = \alpha \psi_2$. Because of this, can define a non degenerate complex product $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbbm{C}$ such that $\mathsf{P}_{\psi_2} (\psi_1) \equiv \frac{\langle \psi_2|\psi_1 \rangle}{|\psi_2|^2} \psi_2$.

@@

Lastly, if we take the projection of $\psi_1$ along $\psi_2$ and we project it again on $\psi_1$, we end up with a component of $\psi_1$. As this is a part of $\psi_1$ prepared as $\psi_1$, its state is just $\psi_1$ rescaled, with no change in correlation. That is: $\mathsf{P}_{\psi_1} (\mathsf{P}_{\psi_2} (\psi_1)) = a \psi_1$ with $a \in \mathbbm{R}$. This tells us that $\langle \psi_2|\psi_1 \rangle = \langle \psi_1|\psi_2 \rangle^\dagger$.

@@

\begin{prop}\label{prop:inner_product}
	The state space $\Psi$ for an irreducible material is an inner product with $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbbm{C}$ such that $\langle \psi_1 | \psi_2 \rangle = \int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) dq^i$.
\end{prop}
\begin{justification}
	We claim there exists a map $| \cdot | ^2 : \Psi \rightarrow \mathbbm{R}$ such that $|\psi|^2 \ge 0 \; \forall \psi \in \Psi$ and $|\psi|^2 = 0 \Leftrightarrow \psi = 0$. Let $| \cdot | ^2 : \Psi \rightarrow \mathbbm{R}$ be the map that returns the amount of material for a given state. Let $\psi \in \Psi$. $|\psi|^2 = \int \rho(q^i) d^nq = \int \psi^\dagger(q^i) \psi(q^i) d^nq \ge 0$. $|\psi|^2 = 0 \Leftrightarrow \psi = 0$ as the only state with no material is the empty state.

	We claim $\forall \psi \in \Psi$ there exist a projection $\mathsf{P}_\psi : \Psi \rightarrow \Psi$ with $\psi \in \Psi$. Let $\psi \in \Psi$ be a state. Let $\mathsf{P}_\psi : \Psi \rightarrow \Psi$ be an operator that returns the component of a given state that is prepared as $\psi$. $\mathsf{P}_{\psi}(c_1 \phi_1 + c_2 \phi_2) = c_1 \mathsf{P}_{\psi}(\phi_1) + c_2 \mathsf{P}_{\psi}(\phi_2)$ is linear as the component of the composition prepared like $\psi$ is the composition of the components so prepared. $\mathsf{P}_\psi \circ \mathsf{P}_\psi = \mathsf{P}_\psi$ is a projection because the result of $\mathsf{P}_\psi$  is prepared as $\psi$.

@@

	We claim $|\psi|^2|\mathsf{P}_{\psi} (\phi)|^2=|\phi|^2|\mathsf{P}_{\phi} (\psi)|^2$. Let $\psi, \phi \in \Psi$. As the material is irreducible, $\psi$ and $\phi$ cannot be used to describe smaller parts. The fraction of $\phi$ that is prepared as $\psi$ must be equal to the fraction of $\psi$ that is prepared as $\phi$. $\frac{|\mathsf{P}_{\psi} (\phi)|^2}{|\phi|^2}=\frac{|\mathsf{P}_{\phi} (\psi)|^2}{|\psi|^2}$.

	We claim $\mathsf{P}_{\psi} (\mathsf{P}_{\phi} (\psi)) = r \psi$ for some $r \in \mathbbm{R}$. Let $\psi, \phi \in \Psi$. Consider $\mathsf{P}_{\psi} (\mathsf{P}_{\phi} (\psi))$: we first take the component of $\psi$ prepared like $\phi$ and then take the component of that prepared like $\psi$. The overall operation is simply returning a part of $\psi$. This must correspond to the same exact configuration for the fragments. The only thing that can change is the amount of material. Therefore $\mathsf{P}_{\psi} (\mathsf{P}_{\phi} (\psi)) = r \psi$ for some $r \in \mathbbm{R}^+$.

@@

	We claim that the inner product can be expressed in terms of the wave functions as $\langle \psi_1 | \psi_2 \rangle = \int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) dq^i$. The amount of material associated with a state $\psi \in \Psi$ is $\langle \psi | \psi \rangle = | \psi |^2 = \int_\mathcal{Q} \psi^\dagger (q^i) \psi(q^i) dq^i$. The inner product associated with the $L^2$ norm is $\int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) dq^i$ with $\psi_1, \psi_2 \in \Psi$. $\langle \psi_1 | \psi_2 \rangle = \int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) dq^i$.

@@

Note that the inner product only has an indirect physical definition, based on the norm and the projections. The former concepts are therefore more useful to gain physical intuition as the latter does not have a straightforward physical meaning.

Also note that the observations made after \ref{prop:differentiable_manifold} for the  classical case are still valid. $\Psi$ cannot be a complete metric space for the norm induced by the inner product as it does not include discontinuous functions. Therefore it is only a pre-Hilbert space. For mathematical convenience, one can take its completion with respect to the norm induced by the inner product and obtain a Hilbert space. Yet, such object does no longer represent a physically meaningful set of states.

Therefore we need to make sure that the rest of our definitions and justifications do not require a complete metric space to be valid or they would not be physically meaningful. In particular, we don't want to use the usual link $\psi(q) = \langle q | \psi \rangle$ between the wave function and the inner product as the material cannot be prepared at a specific value when $q$ is a continuous variable. Similarly, we cannot show that to each state variable is associated an hermitian operator simply by constructing it from its eigenstates as they may not represent physical states.

We proceed in the following way. Suppose we have a fragment variable $f : \mathcal{Q} \rightarrow \mathbbm{R}$. Consider the product $f (q) \psi (q)$: it returns the wave function weighted by the value of $f$. We can think of $f$ as an operator that takes a complex function of $\mathcal{Q}$ and returns another complex function of $\mathcal{Q}$ that represents the distribution of the expectation over the fragments. The operator is linear since $f (q) (a\psi_1 + b\psi_2) = a f \psi_1 + b f  \psi_2$. The operator is hermitian since $f^\dagger = f$. Consider $\int_\mathcal{Q} \psi^\dagger (q) f(q) \psi(q) dq = \int f(q) | \psi(q)|^2 dq$. It integrates the value of $f$ weighted by the amount of the material: it is the expectation of the state variable $f$ over the distribution.

While $\mathcal{Q}$ is enough to define the system of units, it is not enough to define all the fragment variables over which we can define the distributions of irreducible material. The conjugate variables $p_i=\hbar k_i$ cannot be expressed as a function of $q^i$, which reiterates the idea that we cannot specify all variables for each fragment. Suppose there is another maximal set of fragment variables. Then those variables will form a topological space $\hat{\mathcal{Q}}$ such that each $\psi$ can be expressed as a wave function $\psi(\hat{q}) :\hat{\mathcal{Q}} \rightarrow \mathbbm{C}$ over $\hat{\mathcal{Q}}$.

As each state $\psi$ can be identified by a wave function on either space, there must be an operator $\Upsilon^q_{\hat{q}}(\psi(q)) = \psi(\hat{q})$ that converts the wave function on one space to the other.\footnote{Note that a neighbourhood of $q^i$ is not necessarily a neighbourood of $\hat{q}^i$. They are only if $q^i$ and $\hat{q}^i$ can be expressed as a bijection of each other $\hat{q}^i(q^i)$: they are different coordinates of the same manifold.} As to a linear combination of states corresponds a linear combination of wave functions, $\Upsilon^q_{\hat{q}}$ is a linear operator. As converting the space again must give us the original wave function, $\Upsilon^{\hat{q}}_{q} \circ \Upsilon^q_{\hat{q}}= I$ is unitary. We can now define a linear operator $F_{[\hat{q}^i]}$ corresponding to state variable $f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$. Using $\Upsilon^{q}_{\hat{q}}$ we can express $F_{[q^i]} = \Upsilon^{\hat{q}}_{q} \circ f(\hat{q}^i) \Upsilon^{q}_{\hat{q}}$ as an operator over complex functions of $\mathcal{Q}$. It will be still be linear and hermitian.

That is: for each state variable $f$ we have a corresponding linear hermitian operator $F_{[q^i]}$ that acts on wave functions over $q^i$, which can be expressed as function multiplication in a suitable choice of $\hat{\mathcal{Q}}$. The expectation multiplied by the amount of material is given by $\int \psi^\dagger(q^i) F_{[q^i]} (\psi(q^i)) d^nq$.

\begin{prop}\label{prop:self_adjoint_operators}
	For each fragment variable $f$ there exists a topological space $\hat{\mathcal{Q}}$ such that $f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$ and a vector space isomorphism $\Upsilon_{\hat{\mathcal{q}}^i} : \Psi \leftrightarrow \Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ where $\Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ is the space of wave functions over $\hat{\mathcal{Q}}$. There also exists a corresponding self-adjoint linear operator $F_{[q^i]} : \Psi(\mathcal{Q}, dq^i) \rightarrow C(\mathcal{Q}, \mathbbm{C})$ such that $\int \psi^\dagger(q^i) F_{[q^i]}( \psi(q^i))d^nq $ is the expectation of $f$ multiplied by the amount of material. 
\end{prop}
\begin{justification}
	We claim for each fragment variable $f$ there exist a topological space $\hat{\mathcal{Q}}$ such that  $q^i : \mathcal{Q} \rightarrow \mathbbm{R}^n$ provides a maximal description and $f : \mathcal{Q} \rightarrow \mathbbm{R}$. Let $f$ be a fragment variable. Let $\hat{q}^i$ be a set of fragment variables such that $f=f(\hat{q}^i)$. If $\hat{q}^i$ does not provide a maximal set of fragment variables then there exists other fragment variables that refine the description. Extend $\hat{q}^i$ with such variables. If still not maximal, continue extending. As the number of possibilities charted by $\hat{q}^i$ cannot exceed the number of possibilities charted by $q^i$, at some point we'll reach a maximal set. We call $\hat{\mathcal{Q}}$ the topological space charted by the fragment variables $\hat{q}^i$.

	We claim there exists a vector space isomorphism $\Upsilon_{\hat{\mathcal{q}}^i} : \Psi \leftrightarrow \Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$. Let $\psi \in \Psi$ be a state. $\hat{q}^i$ provides a maximal set of fragment variables. There must exist a wave function $\psi(\hat{q}^i) : \mathbbm{R}^n \rightarrow \mathbbm{C}$ associated with $\psi$ such that it gives the maximal description for the state. Let $\Upsilon_{\hat{\mathcal{q}}^i} : \Psi \rightarrow \Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ the map that associates a state with its wave function over $\hat{\mathcal{Q}}$. $\Upsilon_{\hat{\mathcal{q}}^i}$ is a vector space isomorphism for the same reason discussed in \ref{prop:wavefuntion}

	We claim that for each fragment variable $f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$ there exists a corresponding self-adjoint linear operator $F_{[\hat{q}^i]} : \Psi(\hat{\mathcal{Q}}, d\hat{q}^i) \rightarrow C(\hat{\mathcal{Q}}, \mathbbm{C})$ such that $\int \psi^\dagger(\hat{q}^i) F_{[\hat{q}^i]}( \psi(\hat{q}^i)) d^n \hat{q}$ is, if it exists, the expectation of $f$ multiplied by the amount of material. The expectation value of $f$ multiplied by the amount of material is given by $\int f(\hat{q}^i) \rho(\hat{q}^i) d^n\hat{q}$ where $\hat{q}^i$ provide a maximal set of fragment variables such that $f=f(\hat{q})$. We have $\int f(\hat{q}^i) \rho(\hat{q}^i) d^n\hat{q} = \int \psi^\dagger(\hat{q}^i) f(\hat{q}^i) \psi(\hat{q}^i) d^{n}\hat{q}$. Let $F_{[\hat{q}^i]} : \Psi(\hat{\mathcal{Q}}, d\hat{q}^i) \rightarrow C(\hat{\mathcal{Q}}, \mathbbm{C})$ such that $F_{[\hat{q}^i]} (\psi(\hat{q}^i)) = f(\hat{q}^i) \psi(\hat{q}^i)$. $F_{[\hat{q}^i]}$ is a linear operator. $F_{[\hat{q}^i]}$ is self-adjoint as $\int (F_{[\hat{q}^i]}(\psi(\hat{q}^i)))^\dagger \psi(\hat{q}^i) d^{n}\hat{q} = \int (f(\hat{q}^i) \psi(\hat{q}^i))^\dagger \psi(\hat{q}^i) d^{n}\hat{q} = \int \psi^\dagger(\hat{q}^i) f(\hat{q}^i) \psi(\hat{q}^i) d^{n}\hat{q} = \int \psi^\dagger(\hat{q}^i) F_{[\hat{q}^i]} (\psi(\hat{q}^i)) d^{n}\hat{q}$.

	We claim that for any fragment variable $f$ there exists a corresponding self-adjoint linear operator $F_{[q^i]} : \Psi(\mathcal{Q}, dq^i) \rightarrow C(\mathcal{Q}, \mathbbm{C})$ such that $\int \psi^\dagger(q^i) F_{[q^i]}( \psi(q^i)) d^nq$ is, if it exists, the expectation of $f$ multiplied by the amount of material. Let $f$ be a fragment variable. There exist a corresponding self-adjoint linear operator $F_{[\hat{q}^i]} : \Psi(\hat{\mathcal{Q}}, d\hat{q}^i) \rightarrow C(\hat{\mathcal{Q}}, \mathbbm{C})$ where $\hat{\mathcal{Q}}$ provides a maximal set of fragment variables. As both $\hat{\mathcal{Q}}$ and $\mathcal{Q}$ provide a maximal set, we can define $\Upsilon^{q^i}_{\hat{q}^i} \equiv \Upsilon_{\hat{q}^i} \circ \Upsilon_{q^i}^{-1}$ which transforms a wave function over $\mathcal{Q}$ into a wave function over $\hat{\mathcal{Q}}$. $\Upsilon^{q^i}_{\hat{q}^i}$ is an isomorphism as $\Upsilon_{\hat{q}^i}$ and $\Upsilon_{q^i}$ are isomorphisms. $\Upsilon^{q^i}_{\hat{q}^i}$ is unitary as $\Upsilon^{\hat{q}^i}_{q^i} \circ \Upsilon^{q^i}_{\hat{q}^i} = (\Upsilon^{q^i}_{\hat{q}^i})^{-1} \circ \Upsilon^{q^i}_{\hat{q}^i} = I$. We can extend $\Upsilon^{q^i}_{\hat{q}^i}$ on all continuous functions and define $F_{[q^i]} \equiv \Upsilon^{\hat{q}^i}_{q^i} \circ F_{[\hat{q}^i]} \circ \Upsilon^{q^i}_{\hat{q}^i}$. $F_{[q^i]}$ is a self-adjoint linear operator. $\int \psi^\dagger(q^i) F_{[q^i]}( \psi(q^i)) d^nq = \int \psi^\dagger(\hat{q}^i) F_{[\hat{q}^i]}( \psi(\hat{q}^i)) d^n \hat{q}$ is, if it exists, the expectation of $f$ multiplied by the amount of material.

>>>>

For the tool set to be on par with classical mechanics we need to address three requirements. We need to be able to compare two states and ask what fraction of them share the same configuration: this will give us an inner product. We need to be able to express distributions and expectations for different quantities: this will give us self-adjoint linear operators. We then need to identify the self-adjoint linear operators for the quantities $k_i$ and $p_i$ and to be able to express the distribution of material in terms of those. Let's start with the first.

The idea is that, given two states $\psi_1$ and $\psi_2$, we should be able to tell how much of $\psi_1$ is prepared in the same way as $\psi_2$. That is $\psi_1 = \alpha \psi_2 + \psi_3$ is the combination of the component that is prepared like $\psi_2$ (e.g. same distributions in position and momentum) and a component that is prepared completely differently. Let $\mathsf{P}_{\psi_2} : \Psi \rightarrow \Psi$ be the operator that gives us the component $\mathsf{P}_{\psi_2} (\psi_1) \equiv \alpha \psi_2$ prepared like $\psi_2$. This operator is linear in $\psi_1$: the component of the combined system that is prepared like $\psi_2$ is the sum of the subcomponents that are prepared like $\psi_2$. It also gives us the same result if applied once or twice: once we found the component that is prepared like $\psi_2$, it remains the same. $\mathsf{P}_{\psi_2}$ is therefore a projection: a linear operator such that $\mathsf{P}_{\psi_2}(\mathsf{P}_{\psi_2} (\psi_1))= \mathsf{P}_{\psi_2} (\psi_1)$.

If $\psi_1$ is already prepared like $\psi_2$, then its projection on $\psi_2$ will be equal to $\psi_1$. Conversely, if the projection of $\psi_1$ is equal to itself, then it means that the state is already prepared like $\psi_2$. That is: $\mathsf{P}_{\psi_2} (\psi_1) = \psi_1 \Leftrightarrow \psi_1 = \alpha \psi_2$. Because of this, we can define a non degenerate complex product $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbbm{C}$ such that $\mathsf{P}_{\psi_2} (\psi_1) \equiv \frac{\langle \psi_2|\psi_1 \rangle}{|\psi_2|^2} \psi_2$.

@@

Lastly, if we take the projection of $\psi_1$ along $\psi_2$ and we project it again on $\psi_1$, we end up with a component of $\psi_1$. As this is a part of $\psi_1$ prepared as $\psi_1$, its state is just $\psi_1$ rescaled, with no change in correlation. That is: $\mathsf{P}_{\psi_1} (\mathsf{P}_{\psi_2} (\psi_1)) = a \psi_1$ with $a \in \mathbbm{R}^+$. This tells us that $\langle \psi_2|\psi_1 \rangle = \langle \psi_1|\psi_2 \rangle^\dagger$.

@@

\begin{prop}\label{prop:inner_product}
	The state space $\Psi$ for an irreducible material is an inner product with $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbbm{C}$ such that $\langle \psi_1 | \psi_2 \rangle = \int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) d^nq$.
\end{prop}
\begin{justification}
	We claim there exists a map $| \cdot | ^2 : \Psi \rightarrow \mathbbm{R}$ such that $|\psi|^2 \ge 0 \; \forall \psi \in \Psi$ and $|\psi|^2 = 0 \Leftrightarrow \psi = 0$. Let $| \cdot | ^2 : \Psi \rightarrow \mathbbm{R}$ be the map that returns the amount of material for a given state. Let $\psi \in \Psi$ be a state. $|\psi|^2 = \int \rho(q^i) d^nq = \int \psi^\dagger(q^i) \psi(q^i) d^nq \ge 0$. $|\psi|^2 = 0 \Leftrightarrow \psi = 0$ as the only state with no material is the empty state.

	We claim $\forall \psi \in \Psi$ there exists a projection $\mathsf{P}_\psi : \Psi \rightarrow \Psi$ with $\psi \in \Psi$. Let $\psi \in \Psi$ be a state. Let $\mathsf{P}_\psi : \Psi \rightarrow \Psi$ be an operator that returns the component of a given state that is prepared as $\psi$. $\mathsf{P}_{\psi}(c_1 \phi_1 + c_2 \phi_2) = c_1 \mathsf{P}_{\psi}(\phi_1) + c_2 \mathsf{P}_{\psi}(\phi_2)$ is linear as the component of the composition prepared like $\psi$ is the composition of the components so prepared. $\mathsf{P}_\psi \circ \mathsf{P}_\psi = \mathsf{P}_\psi$ is a projection because the result of $\mathsf{P}_\psi$  is prepared as $\psi$.

@@

	We claim $|\psi|^2|\mathsf{P}_{\psi} (\phi)|^2=|\phi|^2|\mathsf{P}_{\phi} (\psi)|^2$. Let $\psi, \phi \in \Psi$. As the material is irreducible, $\psi$ and $\phi$ cannot be used to describe smaller parts. The fraction of $\phi$ that is prepared as $\psi$ must be equal to the fraction of $\psi$ that is prepared as $\phi$. $\frac{|\mathsf{P}_{\psi} (\phi)|^2}{|\phi|^2}=\frac{|\mathsf{P}_{\phi} (\psi)|^2}{|\psi|^2}$. Multiplying by $|\psi|^2|\phi|^2$ we have $|\psi|^2|\mathsf{P}_{\psi} (\phi)|^2=|\phi|^2|\mathsf{P}_{\phi} (\psi)|^2$. This holds for the empty state as well. Let $\phi$ be the empty state. $|\phi|^2=0$ because it has no material and $|\mathsf{P}_{\psi} (\phi)|^2$ as no component of the empty state is prepared like $\psi$.

	We claim $\mathsf{P}_{\psi} (\mathsf{P}_{\phi} (\psi)) = r \psi$ for some $r \in \mathbbm{R}^+$. Let $\psi, \phi \in \Psi$. Consider $\mathsf{P}_{\psi} (\mathsf{P}_{\phi} (\psi))$: we first take the component of $\psi$ prepared like $\phi$ and then take the component of that prepared like $\psi$. The overall operation is simply returning a part of $\psi$. This must correspond to the same exact configuration for the fragments. The only thing that can change is the amount of material. Therefore $\mathsf{P}_{\psi} (\mathsf{P}_{\phi} (\psi)) = r \psi$ for some $r \in \mathbbm{R}^+$.

@@

	We claim that the inner product can be expressed in terms of the wave functions as $\langle \psi_1 | \psi_2 \rangle = \int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) d^nq$. The amount of material associated with a state $\psi \in \Psi$ is $\langle \psi | \psi \rangle = | \psi |^2 = \int_\mathcal{Q} \psi^\dagger (q^i) \psi(q^i) d^nq$. The inner product associated with the $L^2$ norm is $\int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) d^nq$ with $\psi_1, \psi_2 \in \Psi$. $\langle \psi_1 | \psi_2 \rangle = \int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) d^nq$.

@@

Note that the inner product only has an indirect physical definition, based on the norm and the projections. These two concepts are therefore more useful to gain physical intuition as the first does not have a straightforward physical meaning.

Also note that the observations made after \ref{prop:differentiable_manifold} for the  classical case are still valid. $\Psi$ cannot be a complete metric space for the norm induced by the inner product as it does not include discontinuous functions. Therefore it is only a pre-Hilbert space. For mathematical convenience, one can take its completion with respect to the norm induced by the inner product and obtain a Hilbert space. Yet, such object no longer represents a physically meaningful set of states.

Therefore we need to make sure that the rest of our definitions and justifications do not require a complete metric space to be valid or they would not be physically meaningful. In particular, we don't want to use the usual link $\psi(q) = \langle q | \psi \rangle$ between the wave function and the inner product as the material cannot be prepared at a specific value when $q$ is a continuous variable. Similarly, we cannot show that to each state variable is associated a self-adjoint operator $Q = \sum q |q\rangle \langle q |$ simply by expressing it in terms of eigenstates as they may not represent physical states.

We proceed in the following way. Suppose we have a fragment variable $f : \mathcal{Q} \rightarrow \mathbbm{R}$. Consider the product $f (q) \psi (q)$: it returns the wave function weighted by the value of $f$. We can also think of $f$ as an operator that takes a complex function of $\mathcal{Q}$ and returns another complex function of $\mathcal{Q}$ that represents the distribution of the expectation over the fragments. The operator is linear since $f (q) (a\psi_1(q) + b\psi_2(q)) = a f(q) \psi_1(q) + b f(q)  \psi_2(q)$. The operator is self-adjoint since $f^\dagger = f$. Consider $\int_\mathcal{Q} \psi^\dagger (q) f(q) \psi(q) dq = \int f(q) | \psi(q)|^2 dq$. It integrates the value of $f$ weighted by the amount of the material: it is the expectation of the fragment variable $f$ over the distribution.

While $\mathcal{Q}$ is enough to define the system of units, it is not enough to define all the fragment variables over which we can define the distributions of irreducible material. The conjugate variables $p_i=\hbar k_i$ cannot be expressed as a function of $q^i$, which reiterates the idea that we cannot specify all variables for each fragment. Suppose there is another maximal set of fragment variables. Then those variables will form a topological space $\hat{\mathcal{Q}}$ such that each $\psi$ can be expressed as a wave function $\psi(\hat{q}) :\hat{\mathcal{Q}} \rightarrow \mathbbm{C}$.

As each state $\psi$ can be identified by a wave function on either space, there must be an operator $\Upsilon^q_{\hat{q}}(\psi(q)) = \psi(\hat{q})$ that converts the wave function on one space to the other.\footnote{Note that a neighborhood of $q$ is not necessarily a neighborhood of $\hat{q}$. They are only if $q$ and $\hat{q}$ can be expressed as a bijection of each other $\hat{q}^i(q^i)$: they are different coordinates of the same manifold.} As a linear combination of states corresponds to a linear combination of wave functions, $\Upsilon^q_{\hat{q}}$ is a linear operator. As converting the space again must give us the original wave function, $\Upsilon^{\hat{q}}_{q} \circ \Upsilon^q_{\hat{q}}= I$ is unitary. We can now define a linear operator $F_{[\hat{q}]} \equiv f(\hat{q})$ corresponding to fragment variable $f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$. Using $\Upsilon^{q}_{\hat{q}}$ we can express $F_{[q]} = \Upsilon^{\hat{q}}_{q} \circ F_{[\hat{q}]} \circ  \Upsilon^{q}_{\hat{q}}$ as an operator over complex functions of $\mathcal{Q}$. It will still be linear and self-adjoint.

That is: for each fragment variable $f$ we have a corresponding linear self-adjoint operator $F_{[q^i]}$ that acts on wave functions over $q^i$, which can be expressed as function multiplication in a suitable choice of $\hat{\mathcal{Q}}$. The (normalized) expectation of $f$ is given by $\frac{1}{|\psi|^2}\int \psi^\dagger(q^i) F_{[q^i]} (\psi(q^i)) d^nq$.

\begin{prop}\label{prop:self_adjoint_operators}
	For each fragment variable $f$ there exists a topological space $\hat{\mathcal{Q}}$ such that $f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$ and a vector space isomorphism $\Upsilon_{\hat{\mathcal{q}}^i} : \Psi \leftrightarrow \Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ that takes a state $\psi \in \Psi$ and returns a wave function $\psi(\hat{q}^i)$ from $\Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ the space of wave functions over $\hat{\mathcal{Q}}$. There also exists a corresponding self-adjoint linear operator $F_{[q^i]} : \Psi(\mathcal{Q}, dq^i) \rightarrow C(\mathcal{Q}, \mathbbm{C})$ such that $\int \psi^\dagger(q^i) F_{[q^i]}( \psi(q^i))d^nq $ is the expectation of $f$ multiplied by the amount of material. 
\end{prop}
\begin{justification}
	We claim for each fragment variable $f$ there exists a topological space $\hat{\mathcal{Q}}$ such that  $\hat{q}^i : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}^n$ provides a maximal description and $f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$. Let $f$ be a fragment variable. Let $\hat{q}^i$ be a set of fragment variables such that $f=f(\hat{q}^i)$. If $\hat{q}^i$ does not provide a maximal set of fragment variables then there exist other fragment variables that refine the description. Extend $\hat{q}^i$ with such variables. If still not maximal, continue extending. As the number of possibilities charted by $\hat{q}^i$ cannot exceed the number of possibilities charted by $q^i$, at some point we'll reach a maximal set. We call $\hat{\mathcal{Q}}$ the topological space charted by the fragment variables $\hat{q}^i$.

	We claim there exists a vector space isomorphism $\Upsilon_{\hat{\mathcal{q}}^i} : \Psi \leftrightarrow \Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$. Let $\psi \in \Psi$ be a state. $\hat{q}^i$ provides a maximal set of fragment variables. There must exist a wave function $\psi(\hat{q}^i) : \mathbbm{R}^n \rightarrow \mathbbm{C}$ associated with $\psi$ such that it gives the maximal description for the state. Let $\Upsilon_{\hat{\mathcal{q}}^i} : \Psi \rightarrow \Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ be the map that associates a state with its wave function over $\hat{\mathcal{Q}}$. $\Upsilon_{\hat{\mathcal{q}}^i}$ is a vector space isomorphism for the same reason discussed in \ref{prop:wavefuntion}

	We claim that for each fragment variable $f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$ there exists a corresponding self-adjoint linear operator $F_{[\hat{q}^i]} : \Psi(\hat{\mathcal{Q}}, d\hat{q}^i) \rightarrow C(\hat{\mathcal{Q}}, \mathbbm{C})$ such that $\int \psi^\dagger(\hat{q}^i) F_{[\hat{q}^i]}( \psi(\hat{q}^i)) d^n \hat{q}$ is, if it exists, the expectation of $f$ multiplied by the amount of material. The expectation value of $f$ multiplied by the amount of material is given by $\int f(\hat{q}^i) \rho(\hat{q}^i) d^n\hat{q}$ where $\hat{q}^i$ provide a maximal set of fragment variables such that $f=f(\hat{q}^i)$. We have $\int f(\hat{q}^i) \rho(\hat{q}^i) d^n\hat{q} = \int \psi^\dagger(\hat{q}^i) f(\hat{q}^i) \psi(\hat{q}^i) d^{n}\hat{q}$. Let $F_{[\hat{q}^i]} : \Psi(\hat{\mathcal{Q}}, d\hat{q}^i) \rightarrow C(\hat{\mathcal{Q}}, \mathbbm{C})$ such that $F_{[\hat{q}^i]} (\psi(\hat{q}^i)) = f(\hat{q}^i) \psi(\hat{q}^i)$. $F_{[\hat{q}^i]}$ is a linear operator. $F_{[\hat{q}^i]}$ is self-adjoint as $\int (F_{[\hat{q}^i]}(\psi(\hat{q}^i)))^\dagger \psi(\hat{q}^i) d^{n}\hat{q} = \int (f(\hat{q}^i) \psi(\hat{q}^i))^\dagger \psi(\hat{q}^i) d^{n}\hat{q} = \int \psi^\dagger(\hat{q}^i) f(\hat{q}^i) \psi(\hat{q}^i) d^{n}\hat{q} = \int \psi^\dagger(\hat{q}^i) F_{[\hat{q}^i]} (\psi(\hat{q}^i)) d^{n}\hat{q}$.

	We claim that for any fragment variable $f$ there exists a corresponding self-adjoint linear operator $F_{[q^i]} : \Psi(\mathcal{Q}, dq^i) \rightarrow C(\mathcal{Q}, \mathbbm{C})$ such that $\int \psi^\dagger(q^i) F_{[q^i]}( \psi(q^i)) d^nq$ is, if it exists, the expectation of $f$ multiplied by the amount of material. Let $f$ be a fragment variable. There exists a corresponding self-adjoint linear operator $F_{[\hat{q}^i]} : \Psi(\hat{\mathcal{Q}}, d\hat{q}^i) \rightarrow C(\hat{\mathcal{Q}}, \mathbbm{C})$ where $\hat{\mathcal{Q}}$ provides a maximal set of fragment variables. As both $\hat{\mathcal{Q}}$ and $\mathcal{Q}$ provide a maximal set, we can define $\Upsilon^{q^i}_{\hat{q}^i} \equiv \Upsilon_{\hat{q}^i} \circ \Upsilon_{q^i}^{-1}$ which transforms a wave function over $\mathcal{Q}$ into a wave function over $\hat{\mathcal{Q}}$. $\Upsilon^{q^i}_{\hat{q}^i}$ is an isomorphism as $\Upsilon_{\hat{q}^i}$ and $\Upsilon_{q^i}$ are isomorphisms. $\Upsilon^{q^i}_{\hat{q}^i}$ is unitary as $\Upsilon^{\hat{q}^i}_{q^i} \circ \Upsilon^{q^i}_{\hat{q}^i} = (\Upsilon^{q^i}_{\hat{q}^i})^{-1} \circ \Upsilon^{q^i}_{\hat{q}^i} = I$. We can extend $\Upsilon^{q^i}_{\hat{q}^i}$ on all continuous functions and define $F_{[q^i]} \equiv \Upsilon^{\hat{q}^i}_{q^i} \circ F_{[\hat{q}^i]} \circ \Upsilon^{q^i}_{\hat{q}^i}$. $F_{[q^i]}$ is a self-adjoint linear operator. $\int \psi^\dagger(q^i) F_{[q^i]}( \psi(q^i)) d^nq = \int \psi^\dagger(\hat{q}^i) F_{[\hat{q}^i]}( \psi(\hat{q}^i)) d^n \hat{q}$ is, if it exists, the expectation of $f$ multiplied by the amount of material.
